---
title: "Live Session 02 Assignment"
author: "Oscar Padilla"
date: "May 20, 2016"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Calculation Questions
### Use R to calculate the following:
Answers in *italic*

### 1. Basic Math
+ The log of a positive number
```{r}
log(2.718281828)
```
+ What is the default base for the log function? Calculate the log of your previous number with a different base

*The default base for the log function is $e$, thus $log = ln$*
```{r}
log(2.718281828,10)
```
+ The log of a negative number (explain the answer)

*The logarithm of a negative number is undefined*
```{r}
log(-100,10)
```
+ The square-root of a positive number
```{r}
sqrt(25)
```
### 2. Random number generation
+ Create a vector of 15 standard normal random variables. Calculate its mean and SD
```{r}
randomvector <- rnorm(n=15)
randomvector
mean(randomvector)
sd(randomvector)
```
+ Change the mean to 10 and the SD to 2 and recalculate the vector of 15 random normal variables. Calculate its mean and SD
```{r}
randomvector <- rnorm(n=15,mean=10,sd=2)
randomvector
mean(randomvector)
sd(randomvector)
```
+ Why are the means and SD not exactly the same as the means and SDs specified in the function?

*The mean and sd of a set of randomly generated numbers will convey towards the nominal mean and sd as n increases in size*

### 3. Vector Operations
+ The weights of 6 individuals in kg are 60, 72, 57, 90, 95, 72
```{r}
weights <- c(60, 72, 57, 90, 95, 72)
```
+ Their heights (in m) are 1.80, 1.85, 1.72, 1.90, 1.74, 1.91
```{r}
heights <- c(1.80, 1.85, 1.72, 1.90, 1.74, 1.91)
```
+ Enter these vectors into R
```{r}
h_w_table <- data.frame(heights,weights)
h_w_table
```
+ Create a scatterplot of weight vs. height. Interpret the scatterplot.
```{r}
plot(x = heights,y = weights, main= "ScatterPlot Weight v Height", xlab = "m", ylab = "kg")
```

+ Calculate the BMI for each individual (BMI = weight in kg divided by the square of the height in m)
```{r}
BMI <- weights/(heights^2)
BMI
```
+ Calculate the mean for weight
```{r}
avgweight <- mean(weights)
avgweight
```
+ Subtract the mean from each value of weight
```{r}
weights-avgweight
```
+ Sum the result. Now you know why we square the deviations from the mean to calculate a standard deviation!
```{r}
round(sum(weights-avgweight), digits=2)
```
*By definition, the net sum of all deviations from the mean will be zero*